---
title: "7-nmr_markdown"
author: "Kaizad Patel"  
date: "4/7/2020"  
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE,
                      collapse = TRUE,
                      comment = "#>", 
                      fig.path = "images/7b-nmr_markdown/"
                      )

source("0-hysteresis_packages.R")
source("7a-nmr_spectra_setup.R")
source("6b-nmr_peaks_relabund.R")
```

## SPECTRA 
#### spectra
SCL = sandy clay loam (soil)  
SL = sandy loam (soil + sand)  


```{r}
spectra = read.csv("data/processed/nmr_spectra.csv.gz")

spectra_plot = 
  spectra %>% 
  dplyr::mutate(Core = as.character(Core))  %>% 
  na.omit()
```

```{r nmr_spectra_fm, fig.width=15, fig.height=5}
gg_nmr+
  geom_path(data=spectra[spectra$treatment=="FM",], aes(x = ppm, y = intensity, color = as.character(Core)),
            alpha = 0.5)+
  ylim(-0.3,7)+
  facet_grid(.~texture)+theme(legend.position = "none")+
  ggtitle("field moist")
```



```{r nmr_spectra, fig.width=15, fig.height=10}

gg_nmr+
  geom_path(data=spectra_plot[spectra_plot$sat_level=="100",], aes(x = ppm, y = intensity, color = Core),
            alpha = 0.5)+
  ylim(-0.3,7)+
  facet_grid(treatment~texture)+theme(legend.position = "none")+
  ggtitle("100 % saturation")

gg_nmr+
  geom_path(data=spectra_plot[spectra_plot$sat_level=="75",], aes(x = ppm, y = intensity, color = Core),
            alpha = 0.5)+
  ylim(-0.3,7)+
  facet_grid(treatment~texture)+theme(legend.position = "none")+
  ggtitle("75 % saturation")

gg_nmr+
  geom_path(data=spectra_plot[spectra_plot$sat_level=="50",], aes(x = ppm, y = intensity, color = Core),
            alpha = 0.5)+
  ylim(-0.3,7)+
  facet_grid(treatment~texture)+theme(legend.position = "none")+
  ggtitle("50 % saturation")

gg_nmr+
  geom_path(data=spectra_plot[spectra_plot$sat_level=="35",], aes(x = ppm, y = intensity, color = Core),
            alpha = 0.5)+
  ylim(-0.3,7)+
  facet_grid(treatment~texture)+theme(legend.position = "none")+
  ggtitle("35 % saturation")

gg_nmr+
  geom_path(data=spectra_plot[spectra_plot$sat_level=="5",], 
            aes(x = ppm, y = intensity, color = Core),
            alpha = 0.5)+
  ylim(-0.3,7)+
  facet_grid(treatment~texture)+theme(legend.position = "none")+
  ggtitle("5 % saturation")


```

#### representative spectra -- 50 % saturation

```{r nmr_spectra_50_perc, fig.width=8, fig.height=5}
subset_spectra = 
  spectra_plot %>% 
  filter(Core %in% c(120,115))
  
gg_nmr+
  geom_path(data=subset_spectra[subset_spectra$treatment=="Drying",], 
            aes(x = ppm, y = intensity))+
  geom_path(data=subset_spectra[subset_spectra$treatment=="Wetting",], 
            aes(x = ppm, y = intensity+1))+
  annotate("text", label = "rewetting", x = 9, y = 1.1)+
  annotate("text", label = "drying", x = 9, y = 0.1)+
  annotate("text", label = "core 115", x = 10, y = 0.95, size=3, hjust=0)+
  annotate("text", label = "core 120", x = 10, y = -0.05, size=3, hjust=0)+
  
  geom_rect(data = subset_spectra, 
            aes(xmin = DMSO_start, xmax = WATER_stop, ymin = -0.1, ymax = 3.5),
            fill = "grey90", alpha = 0.015)+

  ylim(-0.1,3.5)+
  ylab("")+
  ggtitle("50% saturation")+
  
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
)

```


representative spectra -- 5 % saturation

```{r nmr_spectra_5_perc, fig.width=8, fig.height=5}
subset_spectra_5 = 
  spectra_plot %>% 
  filter(Core %in% c(38,43))
  
gg_nmr+
  geom_path(data=subset_spectra_5[subset_spectra_5$treatment=="Drying",], 
            aes(x = ppm, y = intensity))+
    geom_path(data=subset_spectra_5[subset_spectra_5$treatment=="Wetting",], 
            aes(x = ppm, y = intensity+1))+
  annotate("text", label = "rewetting", x = 9, y = 1.1)+
  annotate("text", label = "drying", x = 9, y = 0.1)+
  annotate("text", label = "core 38", x = 10, y = 0.95, size=3, hjust=0)+
  annotate("text", label = "core 43", x = 10, y = -0.05, size=3, hjust=0)+
  
  geom_rect(data = subset_spectra, 
            aes(xmin = DMSO_start, xmax = WATER_stop, ymin = -0.1, ymax = 3.5),
            fill = "grey90", alpha = 0.015)+

  ylim(-0.1,3.5)+
  ylab("")+
  ggtitle("5% saturation")+
  
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
)

```

## PEAKS
relative abundance for each sample
```{r peaks, fig.width=30, fig.height=30}
ggplot(data=rel_abund[rel_abund$soil_type=="Soil"&!rel_abund$treatment=="FM",], aes(x="", y=relabund, fill=group)) +
  geom_bar(stat="identity", alpha=0.7) +
  #geom_text(aes(x= "", y=pos, label = relabund), size=4) +  # note y = pos
  #facet_wrap(facets = .~source, labeller = label_value) +
  facet_wrap(sat_level+treatment~Core)+
  coord_polar(theta = "y")+
  theme_bw()+
  ggtitle("SCL")


ggplot(data=rel_abund[rel_abund$soil_type=="Soil_sand"&!rel_abund$treatment=="FM",], aes(x="", y=relabund, fill=group)) +
  geom_bar(stat="identity", alpha=0.7) +
  #geom_text(aes(x= "", y=pos, label = relabund), size=4) +  # note y = pos
  #facet_wrap(facets = .~source, labeller = label_value) +
  facet_wrap(sat_level+treatment~Core)+
  coord_polar(theta = "y")+
  theme_bw()+
  ggtitle("SL")

```



## PCA
```{r nmr_pca}
rel_abund = read.csv("data/processed/nmr_rel_abund_cores.csv")


relabund_pca=
  rel_abund %>% 
  ungroup %>% 
  dplyr::select(Core, treatment, sat_level, texture, group, relabund) %>% 
  spread(group, relabund) %>% 
  replace(.,is.na(.),0)  %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM","FM", as.character(sat_level))) %>% 
  dplyr::select(-1)

### scl
relabund_pca_scl_num = 
  relabund_pca %>% 
  filter(texture=="SCL") %>% 
  dplyr::select(.,-(1:3))

relabund_pca_scl_grp = 
  relabund_pca %>% 
  filter(texture=="SCL") %>% 
  dplyr::select(.,(1:3)) %>% 
  dplyr::mutate(row = row_number())

pca_scl = prcomp(relabund_pca_scl_num, scale. = T)
summary(pca_scl)

ggbiplot_scl = 
  ggbiplot(pca_scl, obs.scale = 1, var.scale = 1, 
         groups = relabund_pca_scl_grp$treatment, ellipse = TRUE, circle = F,
         var.axes = TRUE)+
  geom_point(size=2,stroke=2, aes(color = groups, shape = as.factor(relabund_pca_scl_grp$sat_level)))+
  theme_kp()+
  ggtitle("SCL texture")

### sl
relabund_pca_sl_num = 
  relabund_pca %>% 
  filter(texture=="SL") %>% 
  dplyr::select(.,-(1:3))

relabund_pca_sl_grp = 
  relabund_pca %>% 
  filter(texture=="SL") %>% 
  dplyr::select(.,(1:3)) %>% 
  dplyr::mutate(row = row_number())

pca_sl = prcomp(relabund_pca_sl_num, scale. = T)
summary(pca_sl)

ggbiplot_sl = 
  ggbiplot(pca_sl, obs.scale = 1, var.scale = 1, 
         groups = relabund_pca_sl_grp$treatment, ellipse = TRUE, circle = F,
         var.axes = TRUE)+
  geom_point(size=2,stroke=2, aes(color = groups, shape = as.factor(relabund_pca_sl_grp$sat_level)))+
  theme_kp()+
  ggtitle("SL texture")

library(patchwork)
ggbiplot_scl+ggbiplot_sl +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
```


## RELATIVE ABUNDANCE
### tables
```{r}
rel_abund = read.csv("data/processed/nmr_rel_abund.csv")
```

peaks in field moist soils

```{r}
rel_abund %>% 
  filter(treatment == "FM") %>% 
  dplyr::select(group, texture, relative_abundance) %>% 
  spread(texture, relative_abundance) %>% 
  knitr::kable(align = "r")
```

peaks in treatments

```{r}
rel_abund %>% 
  filter(!treatment == "FM") %>% 
  dplyr::mutate(sat_level = factor(sat_level, levels = c(5,35,50,75,100)),
                var = paste0(sat_level,"-",treatment)) %>% 
  dplyr::select(texture, group, var, relative_abundance) %>% 
   
  spread(var, relative_abundance) %>% 
  knitr::kable(align = "r")
```


### rel abund hysteresis?

```{r}
rel_abund_cores = read.csv("data/processed/nmr_rel_abund_cores.csv")
arom = 
  rel_abund_cores %>% 
  filter(group %in% c("aromatic", "aliphatic1", "aliphatic2")) %>% 
  dplyr::select(Core, group, treatment, sat_level, texture, relabund) %>% 
  spread(group, relabund) %>% 
  dplyr::mutate(arom_aliph = aromatic/(aliphatic1+aliphatic2))


ggplot(arom, aes(x = sat_level, y = arom_aliph, color = treatment))+
  geom_point()+
  geom_smooth()+
  ylab("aromatic/aliphatic")+
  facet_grid(.~texture)+
  theme_kp()

ggplot(arom, aes(x = sat_level, y = aromatic, color = treatment))+
  geom_point()+
  geom_smooth()+
  ylab("aromatic")+
  facet_grid(.~texture)+
  theme_kp()
  
```


### rel abund waffle plots

```{r}
library(ggplot2)
library(dplyr)
library(schrute)
library(waffle)
#library(ggtext)
library(cowplot)
library(patchwork)
library(soilpalettes)

rel_abund = read.csv("data/processed/nmr_rel_abund.csv")

rel_abund_50_scl = 
  rel_abund %>% 
  filter(sat_level==50,
         texture=="SCL") %>% 
  ggplot()+
  geom_waffle(aes(fill=group,values=rel_abund),n_rows = 5,size=1.5,color="white", flip = F) + 
  coord_equal() +
scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = soil_palette("podzol",4))+
  #guides(color=F) +
  facet_grid(treatment~.)+
  ggtitle("50% saturation")+
  theme_classic()+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        panel.spacing = unit(1.5, "lines"), #facet spacing for x axis
        strip.background = element_blank(), #facet formatting
        strip.text.y = element_text(size=12, face="bold", angle = 270) #facet labels


)
  
ggdraw(rel_abund_50_scl)
```


all waffle plots
```{r}
rel_abund_scl = 
  rel_abund %>% 
  filter(texture=="SL",
         treatment %in% c("Drying", "Wetting")) %>% 
  ggplot()+
  geom_waffle(aes(fill=group,values=rel_abund),n_rows = 5,size=1.5,color="white", flip = F) + 
  coord_equal() +
scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = soil_palette("podzol"))+
  #guides(color=F) +
  facet_grid(sat_level~treatment)+
  theme_kp()+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank())
  
ggdraw(rel_abund_scl)
```

air-dry waffle plots 

```{r}
rel_abund_scl = 
  rel_abund %>% 
  filter(texture=="SCL",
         sat_level %in% c(5)) %>% 
  ggplot()+
  geom_waffle(aes(fill=group,values=rel_abund),n_rows = 5,size=1.5,color="white", flip = F) + 
  coord_equal() +
scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = soil_palette("podzol"))+
  #guides(color=F) +
  facet_grid(treatment~.)+
  ggtitle("5% saturation") +
  theme_classic()+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        panel.spacing = unit(1.5, "lines"), #facet spacing for x axis
        strip.background = element_blank(), #facet formatting
        strip.text.y = element_text(size=12, face="bold", angle = 270) #facet labels

        )
  
ggdraw(rel_abund_scl)
```

field moist waffle plots 

```{r}
rel_abund_scl_fm = 
  rel_abund %>% 
  filter(texture=="SCL",
         treatment %in% "FM") %>% 
  ggplot()+
  geom_waffle(aes(fill=group,values=rel_abund),n_rows = 5,size=1.5,color="white", flip = F) + 
  coord_equal() +
scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))+
  #scale_fill_manual(values = soil_palette("podzol"))+
  scale_fill_manual(values = c("#351821", "#9F2519", "#FFC87A"))+
  #guides(color=F) +
  facet_grid(treatment~.)+
  ggtitle("field moist") +
  theme_classic()+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        panel.spacing = unit(1.5, "lines"), #facet spacing for x axis
        strip.background = element_blank(), #facet formatting
        strip.text.y = element_text(size=12, face="bold", angle = 270) #facet labels

        )
  
ggdraw(rel_abund_scl_fm)
```

### more viz
```{r relabund_bubble, fig.width=15, fig.height=7}
rel_abund = read.csv("data/processed/nmr_rel_abund.csv")

rel_abund2 = 
  rel_abund %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", "FM", as.character(sat_level)),
                sat_level = factor(sat_level, levels = 
                                     c("FM", "5", "35", "50", "75", "100"))) 
  
  
ggplot(rel_abund2,
       aes(x = treatment, y = group, color = group, size = rel_abund))+
  geom_point()+
  geom_text(data = rel_abund2 %>% 
              filter(
              rel_abund>15),
         aes(label = round(rel_abund,0)), color = "black", size=4)+
  facet_grid(texture ~ sat_level, scales = "free_x")+
  scale_size_area(max_size = 20)+
  theme_bw()


nmr_label <- tribble(
  ~texture, ~sat_level, ~treatment, ~group, ~rel_abund, ~color,
  "SCL", "35", "Wetting", "aliphatic1", 80, "white",
  "SCL", "35", "Wetting", "aromatic", 30, "black",
  "SCL", "50", "Wetting", "aromatic", 30, "black",
  "SCL", "100", "Wetting", "aliphatic2", 80, "white",
  "SL", "35", "Wetting", "aromatic", 30, "black",
  "SL", "50", "Wetting", "aliphatic1", 80, "white",
  "SL", "75", "Wetting", "aromatic", 30, "black",
  "SL", "75", "Wetting", "aliphatic1", 80, "white",
  "SL", "100", "Wetting", "aromatic", 30, "black",
  "SL", "100", "Wetting", "aliphatic1", 80, "white"
) %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", "FM", as.character(sat_level)),
                sat_level = factor(sat_level, levels = 
                                     c("FM", "5", "35", "50", "75", "100"))) 

  


ggplot(rel_abund2,
       aes(x = treatment, y = rel_abund, fill = group))+
  geom_bar(stat = "identity")+
  geom_text(data = nmr_label, 
            aes(x = treatment, y = rel_abund, color = color), 
            label = "*", size=7,
            show.legend = F)+
  facet_grid(texture ~ sat_level, scales = "free_x")+
  scale_x_discrete(position = "top") +
  labs(x = "% saturation")+
  #scale_fill_viridis_d()+
  scale_fill_manual(values = soil_palette("paleustalf",5))+
  scale_color_manual(values = c("black", "white"))+
  theme_bw()
  
  
```








