---
title: "[Hysteresis and Soil C]"
output: github_document
---

# Soil carbon dynamics during drying vs. rewetting: importance of antecedent moisture conditions  
## results

```{r setup, include=FALSE}

knitr::opts_chunk$set(dpi = 300,
                      echo=FALSE,message=FALSE,warning=FALSE,
                      collapse = TRUE,
                      comment = "#>", 
                      fig.path = "images/markdown/")

source("code/0-hysteresis_packages.R")
library(vegan)
```

```{r, eval=TRUE}
pal2 = PNWColors::pnw_palette("Bay",2)
pal3 = c("#dd4124", "#00496f", "#edd746") #bay

  ## pal2=soilpalettes::soil_palette("gley",2)
  ## pal3=c(soilpalettes::soil_palette("gley",2), "black") 
  ##   
  ## pal = soilpalettes::soil_palette("eutrostox",3)
  ## PNWColors::pnw_palette("Bay",3)
  ## pal = c("#ed8b00", "#00496f") 
  ## pal = c("#d2848d", "#6e7cb9") 
  ## pal = c("#f9ad2a", "#1d457f")
```

---

## WATER RETENTION CURVES

```{r hyprop_old, eval=FALSE, fig.width=9, dpi=300, eval=FALSE}
  ##source("2-hyprop.R")
  ##
  ##ggplot(hyprop, aes(y = kPa, x = perc_sat, color = treatment, linetype=treatment))+
  ##  geom_path(size=1.5)+
  ##  scale_color_manual(labels = c("drying","rewetting"), values = c("darkorange2","gray40"),   ##na.translate=F)+
  ##  scale_linetype_manual(labels = c("drying","rewetting"), values = c("solid","twodash"),   ##na.translate=F)+
  ##  facet_wrap(.~soiltype)+
  ##  ylab ("tension, kPa")+
  ##  scale_x_reverse(name = "percent saturation")+
  ##  theme_kp()
  ##
  ##
  ##ggplot(hyprop, aes(y = kPa, x = perc_sat, color = treatment, linetype=treatment))+
  ##  geom_path(size=1.5)+
  ##  scale_color_manual(labels = c("drying","rewetting"), values = c("darkorange2","gray40"),   ##na.translate=F)+
  ##  scale_linetype_manual(labels = c("drying","rewetting"), values = c("solid","twodash"),   ##na.translate=F)+
  ##  facet_wrap(.~soiltype)+
  ##  ylab ("tension, kPa (log scale)")+
  ##  xlab ("percent saturation")+
  ###  scale_x_reverse(name = "percent saturation")+
  ##  scale_y_log10()+
  ## # xlim(0,100)+
  ##  theme_kp()
  ##
  ##
  ##ggplot(hyprop, aes(y = kPa, x = perc_sat, color = treatment, linetype=treatment))+
  ##  geom_path(size=1.5)+
  ##  scale_color_manual(labels = c("drying","rewetting"), values = c("darkorange2","gray40"),   ##na.translate=F)+
  ##  scale_linetype_manual(labels = c("drying","rewetting"), values = c("solid","twodash"),   ##na.translate=F)+
  ##  facet_wrap(.~soiltype)+
  ##  ylab ("tension, kPa (log scale)")+
  ##  xlab ("percent saturation")+
  ###  scale_x_reverse(name = "percent saturation")+
  ##  scale_y_log10()+
  ##  scale_y_continuous(trans = log10_trans(),
  ##                     sec.axis = sec_axis(~ .^-1 * 300, name = "pore size (um)"))+
  ## # xlim(0,100)+
  ##  theme_kp()
  ##

  ##wrc = read.csv("data/processed/wrc.csv")
  ##
  ##ggplot(wrc, aes(y = kPa, x = perc_sat, linetype=treatment, color = treatment))+
  ##  geom_path(size=1)+
  ##  scale_color_manual(labels = c("drying","rewetting"), 
  ##                     values = rev(pal2), 
  ##                     na.translate=F)+
  ##  scale_linetype_manual(labels = c("drying","rewetting"), 
  ##                        values = c("solid","dashed"), 
  ##                        na.translate=F)+
  ##  
  ##  facet_wrap(.~texture)+
  ##  ylab ("tension, kPa")+
  ##  xlab ("percent saturation")+
  ##  scale_y_continuous(trans = scales::log10_trans(), 
  ##                     labels = scales::comma, 
  ##                     sec.axis = sec_axis(~ .^-1 * 300, 
  ##                                         name = "pore size (um)", 
  ##                                         labels = scales::comma))+
  ##  geom_vline(xintercept = 5, size=0.5, color = "grey80")+
  ##  geom_vline(xintercept = 35, size=0.5, color = "grey80")+
  ##  geom_vline(xintercept = 50, size=0.5, color = "grey80")+
  ##  geom_vline(xintercept = 75, size=0.5, color = "grey80")+
  ##  theme_kp()+
  ##  theme(panel.grid = element_blank())+
  ##  NULL
  ##
  ##
  ##

```

```{r wrc_drying, fig.height=5, fig.width=6}
wrc_processed = read.csv("data/processed/wrc.csv")

wrc_processed %>% 
  dplyr::mutate(skip = case_when(grepl("sl_drying",source)~"skip",
                                 grepl("scl_drying",source)~"skip")) %>% 
  filter(is.na(skip)) %>% 
  filter(treatment=="drying") %>% 
  ggplot(aes(x = kPa, y = grav_water_perc, color = texture, group = source))+
  geom_path(size=1.5)+
  scale_color_manual(#labels = c("drying","rewetting"), 
    #values = c("darkorange2","gray40"), 
    values = soilpalettes::soil_palette("eutrostox",2),
    na.translate=F)+
  scale_linetype_manual(labels = c("drying","rewetting"), 
                        values = c("solid","twodash"), na.translate=F)+
  
  labs (x = "tension, kPa (log scale)",
        y = "gravimetric moisture (%)")+
  scale_x_continuous(trans = scales::log10_trans(), labels = scales::comma, 
#                     sec.axis = sec_axis(~ .^-1 * 300, name = "pore size (um)", 
#                                         labels = scales::comma)
)+
            #  geom_hline(yintercept = 5, size=0.5, color = "grey80")+
            #  geom_hline(yintercept = 35, size=0.5, color = "grey80")+
            #  geom_hline(yintercept = 50, size=0.5, color = "grey80")+
            #  geom_hline(yintercept = 75, size=0.5, color = "grey80")+
  theme_kp()+
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, 0.8))+
  NULL

```

```{r wrc_hysteresis, eval=FALSE}
## ggplot for drying and wetting
wrc_processed %>% 
  ggplot(aes(x = kPa, y = grav_water_perc, color = treatment, linetype=treatment, group = source))+
  geom_path(size=1.5)+
  scale_color_manual(labels = c("drying","rewetting"), 
                     values = c("darkorange2","gray40"), na.translate=F)+
  scale_linetype_manual(labels = c("drying","rewetting"), 
                        values = c("solid","twodash"), na.translate=F)+
  
  facet_wrap(.~texture)+
  xlab ("tension, kPa (log scale)")+
  ylab ("gravimetric moisture (%)")+
  scale_x_continuous(trans = scales::log10_trans(), labels = scales::comma, 
                     sec.axis = sec_axis(~ .^-1 * 300, name = "pore size (um)", 
                                         labels = scales::comma))+
  geom_hline(yintercept = 5, size=0.5, color = "grey80")+
  geom_hline(yintercept = 35, size=0.5, color = "grey80")+
  geom_hline(yintercept = 50, size=0.5, color = "grey80")+
  geom_hline(yintercept = 75, size=0.5, color = "grey80")+
  #coord_flip()+
  theme_kp()+
  theme(panel.grid = element_blank())
```

```{r wrc_more, fig.height=6, fig.width=6}
## find closest value ----

wrc_closest = 
  wrc_processed %>% 
  filter(treatment=="drying") %>% 
  group_by(texture) %>% 
  #filter(abs(perc_sat - 50) == min(abs(perc_sat - 50)))
  mutate(points = case_when(
                            abs(perc_sat - 35) == min(abs(perc_sat - 35)) ~ 35,
                            abs(perc_sat - 50) == min(abs(perc_sat - 50)) ~ 50,
                            abs(perc_sat - 75) == min(abs(perc_sat - 75)) ~ 75,
                            abs(perc_sat - 100) == min(abs(perc_sat - 100)) ~ 100),
         poresize_um = round(300/kPa,2)) %>% 
  na.omit()


## labels ----
wrc_label = tribble(
  ~x, ~xend, ~y, ~yend, ~type, ~label, ~label2, ~y2, ~x2,
  30, NA, 175, NA, "text", "100% saturation: \n SCL 3000 μm, SL 3000 μm", "100%: 3000 μm", 100, 10,
  30, NA, 140, NA, "text", "75%: \n SCL 161 μm, SL 186 μm", "75%: 173 μm", 75, 30,
  300, NA, 100, NA, "text", "50%: \n SCL 48 μm, SL 38 μm", "50%: 43 μm", 50, 100,
  300, NA, 70, NA, "text", "35%: \n SCL 17 μm, SL 8 μm", "35%: 12 μm", 35,500,
  
  0.1, 30, 123, 175, "curve", "100%", NA, NA, NA,
  0.1, 30, 185, 175, "curve", "100%", NA, NA, NA,
  
  1.6, 30, 93, 140, "curve", "75%", NA, NA, NA,
  1.9, 30, 140, 140, "curve", "75%", NA, NA, NA,
  
  8, 300, 63, 100, "curve", "50%", NA, NA, NA,
  6, 300, 92, 100, "curve", "50%", NA, NA, NA,
  
  36, 300, 43, 70, "curve", "35%", NA, NA, NA,
  17, 300, 66, 70, "curve", "35%", NA, NA, NA
)

## perc_sat with points ----

wrc_processed %>% 
  dplyr::mutate(skip = case_when(grepl("sl_drying",source)~"skip",
                                 grepl("scl_drying",source)~"skip")) %>% 
  filter(is.na(skip)) %>% 
  filter(treatment=="drying") %>% 
  ggplot(aes(x = kPa, y = perc_sat))+
  geom_path(size=1.5, aes(color = texture, group = source))+
  scale_color_manual(
    values = soilpalettes::soil_palette("eutrostox",2),
    na.translate=F)+
  
  geom_point(data = wrc_closest, size=4, color = "black")+
  geom_text(data = wrc_label %>% filter(type=="text"),
            aes(x = x2, y = y2, label = label2), color = "black")+

  
  labs (x = "tension, kPa (log scale)",
        y = "saturation (%)")+
  scale_x_continuous(trans = scales::log10_trans(), labels = scales::comma, 
                     sec.axis = sec_axis(~ .^-1 * 300, name = "pore size (μm)", 
                                         labels = scales::comma))+
        ##  geom_hline(yintercept = 5, size=0.5, color = "grey80")+
        ##  geom_hline(yintercept = 35, size=0.5, color = "grey80")+
        ##  geom_hline(yintercept = 50, size=0.5, color = "grey80")+
        ##  geom_hline(yintercept = 75, size=0.5, color = "grey80")+
  theme_kp()+
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, 0.8))+
  NULL

## grav with saturation points and pore sizes ----
      ##  wrc_closest_wide = 
      ##    wrc_closest %>% 
      ##    ungroup %>% 
      ##    dplyr::select(texture, points, grav_water_perc, kPa) %>% 
      ##    #spread(texture, grav_water_perc)
      ##    #pivot_wider(from = "kPa")
      ##    nest(kPa, grav_water_perc, .key = 'value_col') %>%
      ##    spread(key = texture, value = value_col) %>%
      ##    unnest(SCL, SL, .sep = '_') %>% 
      ##    mutate(texture="SCL")


## plot ----
wrc_processed %>% 
  dplyr::mutate(skip = case_when(grepl("sl_drying",source)~"skip",
                                 grepl("scl_drying",source)~"skip")) %>% 
  filter(is.na(skip)) %>% 
  filter(treatment=="drying") %>% 
  # geom for wrc ----
  ggplot(aes(x = kPa, y = grav_water_perc))+
  geom_path(size=1.5, aes(color = texture, group = source))+
  scale_color_manual(#labels = c("drying","rewetting"), 
    #values = c("darkorange2","gray40"), 
    values = soilpalettes::soil_palette("eutrostox",2),
    na.translate=F)+
  # geom for labels ----
  geom_text(data = wrc_label %>% filter(type=="text"),
            aes(x = x, y = y, label = label), hjust = "left", vjust = "top")+
  geom_curve(data = wrc_label %>% filter(type=="curve"),
             aes(x  =x, y = y, xend = xend, yend = yend), curvature = -0.3, color = "grey")+  
  # geom for points ----
  geom_point(data = wrc_closest, size=4, color = "black")+


  # labs ----
  labs (x = "tension, kPa (log scale)",
        y = "gravimetric moisture (%)")+
  scale_x_continuous(trans = scales::log10_trans(), labels = scales::comma, 
                     sec.axis = sec_axis(~ .^-1 * 300, name = "pore size (μm)", 
                                         labels = scales::comma))+
  theme_kp()+
  theme(panel.grid = element_blank(),
        legend.position = c(0.1, 0.15))+
  NULL


```

```{r wrc_release}
wrc_release = read.csv("data/processed/wrc_drying_release.csv")

wrc_release %>% 
  mutate(mask = if_else(Diff>5, "large", "small")) %>% 
  ggplot(aes(x = pore_um, y = Diff, color = texture))+
  geom_point()+
  labs(x = "pore size (um)",
       y = "volume of pores (cm3)")+
  scale_color_manual(values = soilpalettes::soil_palette("eutrostox",2))+
  #facet_grid(mask~., scales = "free_y")+
  scale_y_log10(breaks=c(0.1, 0.2, 0.5, 1, 5, 10))+
  theme_kp()+
  theme(panel.grid.minor = element_blank())+
  NULL

release_drying_summary  = 
  wrc_release %>% 
  group_by(start, stop, treatment, texture) %>% 
  summarize(Diff = sum(Diff), # cm3
            Diff_perc = sum(Diff_perc)) 

release_drying_summary %>% 
  ggplot(aes(x = stop, y = Diff, color = texture))+
  geom_point()+
  labs(x = "pore size (um)",
       y = "volume of pores (cm3)",
       subtitle = "pore size in 10 um bins")+
  scale_color_manual(values = soilpalettes::soil_palette("eutrostox",2))+
  #facet_grid(mask~., scales = "free_y")+
  scale_y_log10(breaks=c(0.1, 0.2, 0.5, 1, 5, 10))+
  scale_x_log10()+
  theme_kp()+
  theme(panel.grid.minor = element_blank())+
  NULL
```

```{r wrc_release_bins, eval=FALSE}
release_drying_bins = 
  release_drying_summary %>% 
  ungroup() %>% 
  arrange(desc(stop)) %>% 
  mutate(bins = if_else(stop<=100, stop, 3000)) %>% 
  group_by(treatment, texture, bins) %>% 
  dplyr::summarise(vol_water_cm3 = sum(Diff),
                   vol_water_perc = sum(Diff_perc))

release_drying_bins %>% 
  ggplot(aes(x = reorder(as.character(bins), bins), y = vol_water_cm3, color = texture))+
  geom_point(size=3.5)+
  labs(x = "pore size class, upper limit (um)",
       y = "volume of pores in size class (cm3)")+
  scale_color_manual(values = soilpalettes::soil_palette("eutrostox",2))+
  theme_kp()+
  NULL
```


---

## RESPIRATION

```{r picarro_files}
fluxes = read.csv("data/processed/picarro_fluxes.csv", stringsAsFactors = F) %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", as.integer(53), sat_level))

meanfluxes_summary = read.csv("data/processed/picarro_meanfluxes_summary.csv", stringsAsFactors = F)
```

```{r picarro_anova}
aov2 = aov(flux_co2_nmol_gC_s ~ treatment*texture*sat_level, data = fluxes)
summary(aov2)
```

```{r respiration_graph, fig.width=9}

label = tibble(x = c(50, 75, 100, 100),
               y = c(0.80, 1.10, 1.10, 0.90),
               texture = c("SCL", "SCL", "SCL", "SL"),
               label = c("*","*","*","*"),
               treatment = c("Drying", "Drying", "Drying", "Drying"))

ggplot(fluxes, aes(x = sat_level, y = flux_co2_nmol_gC_s, shape=treatment, color = treatment))+
  geom_point(position = position_dodge(width = 0.5), size=3.5, stroke=1.5)+
  geom_text(data = label, aes(x = x, y = y, label = label), size=10, color = "black")+
  scale_color_manual(values = (pal3),
                     breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"), 
                     na.translate=F)+
  scale_shape_manual(breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"), 
                     values = c(16,1, 4), na.translate=F)+
  facet_grid(.~texture)+
  ylab(expression(bold(paste("mean CO"[2]," flux, nmol g"^{-1}, "C s"^{-1}))))+
  xlab("percent saturation")+
  theme_kp()+
  NULL
```

---

## WSOC
```{r wsoc_files}
wsoc = read.csv(WSOC, stringsAsFactors = F) %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", as.integer(53), sat_level))
wsoc_summary = read.csv("data/processed/wsoc_summary.csv", stringsAsFactors = F)
wsoc_label =
  wsoc %>%
  group_by(texture, sat_level) %>% 
  dplyr::mutate(y = max(wsoc_mg_g)+0.05) %>% 
  ungroup() %>% 
  mutate(asterisk = case_when(texture=="SCL" & sat_level %in% c(5,35,50,75,100) ~ "*",
                              texture=="SL" & sat_level %in% c(5,50,75,100) ~ "*")) %>% 
  na.omit()
```

```{r wsoc_graphs, fig.width=9}
ggplot(wsoc, aes(x = sat_level, y = wsoc_mg_g, shape = treatment, color = treatment))+
  geom_point(size=3.5, stroke=1.5)+
  geom_text(data = wsoc_label %>% na.omit(), 
            aes(x = sat_level, y = y, label = asterisk), color = "black", size=10)+
  scale_color_manual(values = pal3,
                     breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"), 
                    na.translate=F)+
  scale_shape_manual(breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"), 
                     values = c(16,1,4), na.translate=F)+
  facet_grid(.~texture)+
  ylab(expression(bold(paste("WEOC, mg gC"^{-1}))))+
  xlab("percent saturation")+
  theme_kp()+
  NULL
```

```{r wsoc_anova}
aov = aov(wsoc_mg_g ~ treatment * texture * sat_level, data = wsoc[!wsoc$sat_level==5,])
summary(aov)

```

For field moist conditions, WSOC was significantly greater in SL than in SCL soils. 

```{r, echo=TRUE}
aov_wsoc_fm = aov(wsoc_mg_g ~ texture, data = wsoc %>% filter(treatment=="FM"))
summary(aov_wsoc_fm)
h = agricolae::HSD.test(aov_wsoc_fm, trt = "texture");h
```


---

## FTICR

```{r}
fticr_data = read.csv(FTICR_LONG)
#fticr_meta = read.csv(FTICR_META)
meta_hcoc = read.csv(FTICR_META_HCOC)

#relabund = read.csv("data/processed/fticr_relabund.csv")
corekey = read.csv(COREKEY)

corekey_subset = 
  corekey %>% 
  dplyr::select(Core, texture, treatment, perc_sat, sat_level,Core_assignment) %>% 
  dplyr::mutate(Core = as.factor(Core))

fticr = 
  fticr_data %>% 
  dplyr::mutate(sat_level = if_else(treatment == "FM", 50, as.numeric(sat_level))) %>% 
  left_join(meta_hcoc, by = "formula")
```

```{r vk2, fig.height = 6, fig.width=10}
fticr_subset = 
  fticr %>% 
  filter(sat_level %in% c(50,100)) %>% 
  filter(!(sat_level==50 & (treatment == "Drying"|treatment=="Wetting")))

gg_fm = gg_vankrev(fticr_subset[fticr_subset$treatment=="FM",], aes(x = OC, y = HC, color = treatment))+
  scale_color_manual(values = "#edd746")+
  annotate("text", label = "aliphatic", x = 0.2, y = 1.75)+
   annotate("text", label = "highly unsaturated/ \n lignin-like", x = 0.2, y = 1.25)+
   annotate("text", label = "aromatic", x = 0.2, y = 0.85)+
   annotate("text", label = "condensed aromatic", x = 0.2, y = 0.35)+
  theme_kp()+
  theme(legend.position = c(0.7, 0.1))

gg_100 = 
  fticr_subset %>% 
  filter(sat_level==100) %>% 
  #dplyr::mutate(treatment = factor(treatment, levels = c("Wetting", "Drying"))) %>% 
  gg_vankrev( 
             aes(x = OC, y = HC, color = treatment, group = rev(treatment)))+
  scale_color_manual(values = rev(pal2))+
  #geom_point(size=0.5, alpha = 0.4)+
  #facet_grid(.~treatment)+
  theme_kp()+
  theme(legend.position = c(0.7, 0.1),
        legend.text.align = 1)+
  guides(colour = guide_legend(override.aes = list(alpha=1),nrow = 1))
```

```{r vk_marginal, fig.height=5, fig.width=10}

gg_marg_fm = ggMarginal(gg_fm,groupColour = TRUE,groupFill = TRUE)
gg_marg_100 = ggMarginal(gg_100,groupColour = TRUE,groupFill = TRUE)

plot_grid(gg_marg_fm, gg_marg_100, nrow = 1, ncol = 2)
```

### PERMANOVA

```{r}
relabund = read.csv("data/processed/fticr_relabund.csv")

# 1. PERMANOVA ----

## make wider 
relabund2 = 
  relabund %>% 
  dplyr::select(Core, texture, treatment, sat_level, Class, relabund) %>% 
  spread(Class, relabund) %>% 
  filter(!treatment=="FM") %>% 
  filter(texture=="SCL") %>% 
  replace(is.na(.),0)

# create a matrix within relabund2 file, which we will use as the PERMANOVA response variable
relabund2$DV = as.matrix(relabund2[,5:8])

# PERMANOVA
adonis2(relabund2$DV ~ treatment*sat_level, data = relabund2)
```


### relative abundance bar plots

```{r fticr_barplot, fig.width=9, fig.height=5}
library(scales)

relabund_trt = read.csv("data/processed/fticr_relabund_trt.csv")


fticr_label <- 
  tribble(
  ~texture, ~sat_level, ~treatment, ~Class, ~rel_abund, ~color,
  "SCL", "5", "Wetting", "aliphatic", 90, "white",
  "SCL", "5", "Wetting", "condensed aromatic", 1, "black",
  
  "SCL", "35", "Wetting", "aliphatic", 90, "white",
  "SCL", "35", "Wetting", "unsaturated/lignin", 50, "black",
  "SCL", "35", "Wetting", "aromatic", 10, "black",
  
  "SCL", "50", "Wetting", "aliphatic", 90, "white",
  "SCL", "50", "Wetting", "unsaturated/lignin", 50, "black",
  "SCL", "50", "Wetting", "aromatic", 10, "black",
  "SCL", "50", "Wetting", "condensed aromatic", 1, "black",
  
  "SCL", "75", "Wetting", "aliphatic", 90, "white",
  "SCL", "75", "Wetting", "unsaturated/lignin", 50, "black",
  "SCL", "75", "Wetting", "aromatic", 10, "black",
  
  "SCL", "100", "Wetting", "condensed aromatic", 1, "black"
) %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", "FM", as.character(sat_level)),
                sat_level = factor(sat_level, levels = 
                                     c("FM", "5", "35", "50", "75", "100"))) 


relabund_trt %>% 
  filter(texture=="SCL") %>% 
  ungroup %>% 
  mutate(Class = factor(Class, levels = 
                          c("aliphatic",
                            "unsaturated/lignin",
                            "aromatic",
                            "condensed aromatic"))) %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", "FM", as.character(sat_level)),
                sat_level = factor(sat_level, levels = 
                                     c("FM", "5", "35", "50", "75", "100"))) %>% 
  ggplot(aes(x = treatment, y = rel_abund, fill = Class))+
  geom_bar(stat = "identity")+
  
  geom_text(data = fticr_label, 
            aes(x = treatment, y = rel_abund, color = color), 
            label = "*", size=7,
            show.legend = F)+
  
  scale_fill_manual(values = PNWColors::pnw_palette("Sunset2",4))+
  scale_color_manual(values = c("black", "white"))+

  labs(x = "",
       y = "relative contribution (%)",
       subtitle = "% saturation")+
  #scale_y_continuous(limits=c(60,100), oob = rescale_none)+
  facet_grid(texture~sat_level, scales = "free_x", space = "free_x")+
  theme_kp()+
  theme(
    plot.subtitle = element_text(size=14, face = "bold", hjust=0.5),
    panel.border = element_rect(color="black",size=0.5, fill = NA),
    strip.background = element_blank(), #facet formatting
    panel.spacing.x = unit(0.5, "lines"), #facet spacing for x axis
    panel.spacing.y = unit(1, "lines"), #facet spacing for x axis
    strip.text.x = element_text(size=12, face = "bold"), #facet labels
    strip.text.y = element_text(size=12, face = "bold", angle = 270)
    )+
  NULL
```

### relative abundance PCA

```{r fticr_pca}
library(scales)
relabund = read.csv("data/processed/fticr_relabund.csv")

# PART II: RELATIVE ABUNDANCE PCA ----
relabund_pca=
  relabund %>% 
  ungroup %>% 
  filter(texture=="SCL") %>% 
  #filter(!Core==25) %>% 
  dplyr::select(Core, Core_assignment, treatment, sat_level, texture, Class, relabund) %>% 
  spread(Class, relabund) %>% 
  replace(.,is.na(.),0)  %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM","FM", as.character(sat_level))) %>% 
  dplyr::select(-1)

#
## SCL PCA ----
relabund_pca_num = 
  relabund_pca %>% 
  dplyr::select(.,-(1:4))

relabund_pca_grp = 
  relabund_pca %>% 
  dplyr::select(.,(1:4)) %>% 
  dplyr::mutate(row = row_number())

pca = prcomp(relabund_pca_num, scale. = T)
#summary(pca)


## biplot ----
fticr_biplot = 
  ggbiplot(pca, obs.scale = 1, var.scale = 1, 
         groups = relabund_pca_grp$treatment, ellipse = TRUE, circle = F,
         var.axes = TRUE)+
  geom_point(size=2,stroke=2, 
             aes(color = groups, 
                 shape = as.factor(relabund_pca_grp$sat_level)))+
  scale_color_manual(values = (pal3),
                     breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"),
                     na.translate=F) +
  scale_shape_manual(values = c(18, 17, 3, 15, 16, 4),
                     breaks = c(5, 35, 50, 75, 100, "FM"))+
  scale_x_continuous(limits = c(-3.5,3.5), oob=rescale_none)+
  scale_y_continuous(limits = c(-3,3), oob=rescale_none)+
  #labs(title = "SCL")+
  theme_kp()+
  theme(panel.grid = element_blank(),
        title = element_text(hjust = 0.5),
        legend.position = c(0, 1),
        legend.direction="vertical",
        legend.justification = c(0,0.8),

        legend.text = element_text(size = 10),
        legend.spacing.y = unit(0, 'mm'))+
  guides(shape = guide_legend(ncol = 2))+
  NULL
```

```{r fticr_eucl}
# euclidean distance ----
library(vegan)
library(ape)

# make matrices ----
bray_distance = vegdist(relabund_pca_num, method="euclidean")
principal_coordinates = pcoa(bray_distance)
pcoa_plot = data.frame(principal_coordinates$vectors[,])
pcoa_plot_merged = merge(pcoa_plot, relabund_pca_grp, by="row.names")

grp = 
  relabund_pca_grp %>% 
  dplyr::mutate(grp = paste0(texture,"-",sat_level,"-",treatment))
  #dplyr::select(row, grp)
matrix = as.matrix(bray_distance)

matrix2 = 
  matrix %>% 
  melt() %>% 
  left_join(grp, by = c("Var1"="row")) %>% 
  #rename(grp1 = grp) %>% 
  left_join(grp, by = c("Var2"="row")) %>% 
  filter(grp.x==grp.y) %>% 
  group_by(grp.x,grp.y,sat_level.x, texture.x,treatment.x,treatment.y) %>% 
  dplyr::summarise(distance  =mean(value)) %>%
  ungroup %>% 
  rename(sat_level = sat_level.x) %>% 
  dplyr::mutate(sat_level = if_else(treatment.x=="FM","FM",sat_level),
                sat_level = factor(sat_level, levels = c(5,35,50,75,100,"FM")))

matrix3 = 
  matrix %>% 
  melt() %>% 
  left_join(grp, by = c("Var1"="row")) %>% 
  #rename(grp1 = grp) %>% 
  left_join(grp, by = c("Var2"="row")) %>% 
  filter(!grp.x==grp.y) %>% 
  filter(sat_level.x == sat_level.y) %>% 
  filter(texture.x == texture.y) %>% 
  group_by(grp.x,grp.y,sat_level.x, texture.x,treatment.x,treatment.y) %>% 
  dplyr::summarise(distance  =mean(value)) %>%
  ungroup %>% 
  rename(sat_level = sat_level.x) %>% 
  dplyr::mutate(sat_level = factor(sat_level, levels = c(5,35,50,75,100)))


eucl_label = tribble(
  ~x, ~xend, ~y, ~yend, ~texture.x, ~type,
  0.7, NA , 3, NA, "SCL", "text",
  0.7, 0.7, 2, 2.8, "SCL", "curve"
)
  
# make plot ----
fticr_eucl = 
  ggplot(matrix3, aes(x = sat_level, y = distance))+
  geom_point(size=3)+
  geom_segment(aes(x = sat_level, xend = sat_level, y = 0, yend = distance))+
  
  ylim(0,10)+
  labs(x = "% saturation",
       y = "drying-rewetting \n Euclidean distance")+
  geom_hline(yintercept = 1.713, linetype = "dashed")+
  
##  geom_text(data = eucl_label %>% filter(type=="text"), aes(x = x, y = y), 
##            label = "avg. within-group distance", angle=90, hjust="left", size=3)+
##  geom_curve(data = eucl_label %>% filter(type=="curve"), 
##             aes(x = x, xend = xend, y = y, yend = yend),
##             curvature = -0.2)+
  
  theme_kp()

  # within-treatment variation
  # ggplot(matrix2, aes(x = sat_level, y = mean(distance), color = treatment.x, shape = texture.x))+
  #   geom_point(size = 3)+
  #   ylim(1,2)+
  #   theme_kp()
# mean(matrix2$distance)

```

```{r fticr_pca_combined, fig.width=7, fig.height=9}
library(ggpubr)
ggarrange(fticr_biplot , fticr_eucl,
          common.legend = TRUE, legend="right",
          nrow = 2,
          heights = c(1.5,0.8), align = "hv",
          labels = c("a", "b")
)
```


---

## NMR

### spectra 
```{r, eval=TRUE}
spectra = read.csv("data/processed/nmr_spectra.csv.gz")
source("code/7a-nmr_spectra_setup.R")

spectra_plot = 
  spectra %>% 
  dplyr::mutate(Core = as.character(Core))  %>% 
  na.omit()
```

#### representative spectra -- 50 % saturation

```{r nmr_spectra_50_perc, fig.width=8, fig.height=5, eval=TRUE}
subset_spectra = 
  spectra_plot %>% 
  filter(Core %in% c(120,115))
  
gg_nmr+
  geom_path(data=subset_spectra[subset_spectra$treatment=="Drying",], 
            aes(x = ppm, y = intensity))+
  geom_path(data=subset_spectra[subset_spectra$treatment=="Wetting",], 
            aes(x = ppm, y = intensity+1))+
  annotate("text", label = "rewetting", x = 9, y = 1.1)+
  annotate("text", label = "drying", x = 9, y = 0.1)+
  annotate("text", label = "core 115", x = 10, y = 0.95, size=3, hjust=0)+
  annotate("text", label = "core 120", x = 10, y = -0.05, size=3, hjust=0)+
  
  geom_rect(data = subset_spectra, 
            aes(xmin = DMSO_start, xmax = WATER_stop, ymin = -0.1, ymax = 3.5),
            fill = "grey90", alpha = 0.015)+

  ylim(-0.1,3.5)+
  ylab("")+
  annotate("text", label = "50% saturated", x = 9, y = 3.5)+

  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
)
```

representative spectra -- 5 % saturation

```{r nmr_spectra_5_perc, fig.width=8, fig.height=5, eval=FALSE}
subset_spectra_5 = 
  spectra_plot %>% 
  filter(Core %in% c(38,43))
  
gg_nmr+
  geom_path(data=subset_spectra_5[subset_spectra_5$treatment=="Drying",], 
            aes(x = ppm, y = intensity))+
    geom_path(data=subset_spectra_5[subset_spectra_5$treatment=="Wetting",], 
            aes(x = ppm, y = intensity+1))+
  annotate("text", label = "rewetting", x = 9, y = 1.1)+
  annotate("text", label = "drying", x = 9, y = 0.1)+
  annotate("text", label = "core 38", x = 10, y = 0.95, size=3, hjust=0)+
  annotate("text", label = "core 43", x = 10, y = -0.05, size=3, hjust=0)+
  
  geom_rect(data = subset_spectra, 
            aes(xmin = DMSO_start, xmax = WATER_stop, ymin = -0.1, ymax = 3.5),
            fill = "grey90", alpha = 0.015)+

  ylim(-0.1,3.5)+
  ylab("")+
  ggtitle("5% saturation")+
  
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
)

```


### PCA
```{r nmr_pca}
rel_abund_cores = read.csv("data/processed/nmr_rel_abund_cores.csv")

# pca ----
relabund_pca=
  rel_abund_cores %>% 
  ungroup %>% 
  dplyr::select(Core, treatment, sat_level, texture, group, relabund) %>% 
  spread(group, relabund) %>% 
  replace(.,is.na(.),0)  %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM","FM", as.character(sat_level))) %>% 
  dplyr::select(-1)

# pca scl ----
relabund_pca_scl_num = 
  relabund_pca %>% 
  filter(texture=="SCL") %>% 
  dplyr::select(.,-(1:3))

relabund_pca_scl_grp = 
  relabund_pca %>% 
  filter(texture=="SCL") %>% 
  dplyr::select(.,(1:3)) %>% 
  dplyr::mutate(row = row_number())

pca_scl = prcomp(relabund_pca_scl_num, scale. = T)
#summary(pca_scl)

## biplot scl ----
ggbiplot_scl = 
    ggbiplot(pca_scl, obs.scale = 1, var.scale = 1, 
         groups = relabund_pca_scl_grp$treatment, ellipse = TRUE, circle = F,
         var.axes = TRUE,
         )+
    geom_point(size=2,stroke=2, 
               aes(color = groups,
                   shape = as.factor(relabund_pca_scl_grp$sat_level)))+
    scale_color_manual(values = (pal3),
                     breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"), 
                     na.translate=F)+
    scale_shape_manual(values = c(18, 17, 3, 15, 16, 4),
                       breaks = c(5, 35, 50, 75, 100, "FM"))+
    ggtitle("SCL")+
    theme_kp()+
    theme(panel.grid = element_blank(),
          title = element_text(hjust = 0.5))+
    NULL


# pca sl ----
relabund_pca_sl_num = 
  relabund_pca %>% 
  filter(texture=="SL") %>% 
  dplyr::select(.,-(1:3))

relabund_pca_sl_grp = 
  relabund_pca %>% 
  filter(texture=="SL") %>% 
  dplyr::select(.,(1:3)) %>% 
  dplyr::mutate(row = row_number())

pca_sl = prcomp(relabund_pca_sl_num, scale. = T)
#summary(pca_sl)

## biplot sl ----
ggbiplot_sl = 
  ggbiplot(pca_sl, obs.scale = 1, var.scale = 1,
           groups = relabund_pca_sl_grp$treatment, ellipse = TRUE, circle = F,
           var.axes = TRUE)+
   geom_point(size=2,stroke=2, 
              aes(color = groups,
                  shape = as.factor(relabund_pca_sl_grp$sat_level)))+
    scale_color_manual(values = (pal3),
                     breaks = c("Drying","Wetting", "FM"),
                     labels = c("drying","rewetting", "field moist"), 
                     na.translate=F)+
    scale_shape_manual(values = c(18, 17, 3, 15, 16, 4),
                       breaks = c(5, 35, 50, 75, 100, "FM"))+
  ggtitle("SL")+
    theme_kp()+
    theme(panel.grid = element_blank(),
          title = element_text(hjust = 0.5))+  
  NULL


# combined biplots ----
library(patchwork)
ggbiplot_scl+ggbiplot_sl +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r nmr_euclidean, fig.height=3.5, fig.width=6.5}
## 1. overall PCA ----
relabund_pca_num = 
  relabund_pca %>% 
  dplyr::select(.,-(1:3))

relabund_pca_grp = 
  relabund_pca %>% 
  dplyr::select(.,(1:3)) %>% 
  dplyr::mutate(row = row_number())

# euclidean distance ----
library(vegan)
library(ape)

## matrices ----
bray_distance = vegdist(relabund_pca_num, method="euclidean")
principal_coordinates = pcoa(bray_distance)
pcoa_plot = data.frame(principal_coordinates$vectors[,])
pcoa_plot_merged = merge(pcoa_plot, relabund_pca_grp, by="row.names")

grp = 
  relabund_pca_grp %>% 
  dplyr::mutate(grp = paste0(texture,"-",sat_level,"-",treatment))
  #dplyr::select(row, grp)
matrix = as.matrix(bray_distance)

matrix2 = 
  matrix %>% 
  melt() %>% 
  left_join(grp, by = c("Var1"="row")) %>% 
  #rename(grp1 = grp) %>% 
  left_join(grp, by = c("Var2"="row")) %>% 
  filter(grp.x==grp.y) %>% 
  group_by(grp.x,grp.y,sat_level.x, texture.x,treatment.x,treatment.y) %>% 
  dplyr::summarise(distance  =mean(value)) %>%
  ungroup %>% 
  rename(sat_level = sat_level.x) %>% 
  dplyr::mutate(sat_level = if_else(treatment.x=="FM","FM",sat_level),
                sat_level = factor(sat_level, levels = c(5,35,50,75,100,"FM")))

matrix3 = 
  matrix %>% 
  melt() %>% 
  left_join(grp, by = c("Var1"="row")) %>% 
  #rename(grp1 = grp) %>% 
  left_join(grp, by = c("Var2"="row")) %>% 
  filter(!grp.x==grp.y) %>% 
  filter(sat_level.x == sat_level.y) %>% 
  filter(texture.x == texture.y) %>% 
  group_by(grp.x,grp.y,sat_level.x, texture.x,treatment.x,treatment.y) %>% 
  dplyr::summarise(distance  =mean(value)) %>%
  ungroup %>% 
  rename(sat_level = sat_level.x) %>% 
  dplyr::mutate(sat_level = factor(sat_level, levels = c(5,35,50,75,100)))

## plot ----
eucl_label = tribble(
  ~x, ~xend, ~y, ~yend, ~texture.x, ~type,
  0.7, NA , 30, NA, "SCL", "text",
  0.7, 0.7, 19, 28, "SCL", "curve"
)
  
ggplot(matrix3, aes(x = sat_level, y = distance))+
  geom_point(size=3)+
  geom_segment(aes(x = sat_level, xend = sat_level, y = 0, yend = distance))+
  ylim(0,80)+
  labs(x = "% saturation",
       y = "drying-rewetting \n Euclidean distance")+
  geom_hline(yintercept = 17.52, linetype = "dashed")+
  
  geom_text(data = eucl_label %>% filter(type=="text"), aes(x = x, y = y), 
            label = "avg. within-group distance", angle=90, hjust="left", size=3)+
  geom_curve(data = eucl_label %>% filter(type=="curve"), 
             aes(x = x, xend = xend, y = y, yend = yend),
             curvature = -0.2)+
  
  facet_grid(.~texture.x)+
  theme_kp()

  # within-treatment variation
  # ggplot(matrix2, aes(x = sat_level, y = mean(distance), color = treatment.x, shape = texture.x))+
  #   geom_point(size = 3)+
  #   ylim(0,80)+
  #   theme_kp()
# mean(matrix2$distance)
```

### RELATIVE ABUNDANCE

```{r}
rel_abund = read.csv("data/processed/nmr_rel_abund.csv")
rel_abund_cores = read.csv("data/processed/nmr_rel_abund_cores.csv")

```

### stats
PERMANOVA: treatment and moisture level had a significant influence

```{r nmr_permanova}
rel_abund2 = 
  rel_abund_cores %>% 
  dplyr::select(Core, texture, treatment, sat_level, group, relabund) %>% 
  spread(group, relabund) %>% 
  replace(is.na(.),0) 

rel_abund2$DV = as.matrix(rel_abund2[,5:9])

library(vegan)
adonis2(rel_abund2$DV ~ treatment*sat_level +sat_level*texture + treatment*texture, data = rel_abund2)

adonis2(rel_abund2$DV ~ treatment*sat_level, data = rel_abund2)

```



### tables

peaks in field moist soils

```{r}
rel_abund %>% 
  filter(treatment == "FM") %>% 
  dplyr::select(group, texture, relative_abundance) %>% 
  spread(texture, relative_abundance) %>% 
  knitr::kable(align = "r")
```

peaks in treatments

```{r}
rel_abund %>% 
  filter(!treatment == "FM") %>% 
  dplyr::mutate(sat_level = factor(sat_level, levels = c(5,35,50,75,100)),
                var = paste0(sat_level,"-",treatment)) %>% 
  dplyr::select(texture, group, var, relative_abundance) %>% 
   
  spread(var, relative_abundance) %>% 
  knitr::kable(align = "r")
```



### more viz
```{r nmr_relabund, fig.width=9, fig.height=7}
rel_abund2 = 
  rel_abund %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", "FM", as.character(sat_level)),
                sat_level = factor(sat_level, levels = 
                                     c("FM", "5", "35", "50", "75", "100"))) 

nmr_label <- tribble(
  ~texture, ~sat_level, ~treatment, ~group, ~rel_abund, ~color,
  "SCL", "35", "Wetting", "aliphatic1", 80, "white",
  "SCL", "35", "Wetting", "aromatic", 30, "black",
  "SCL", "50", "Wetting", "aromatic", 30, "black",
  "SCL", "100", "Wetting", "aliphatic2", 80, "white",
  "SL", "35", "Wetting", "aromatic", 30, "black",
  "SL", "50", "Wetting", "aliphatic1", 80, "white",
  "SL", "75", "Wetting", "aromatic", 30, "black",
  "SL", "75", "Wetting", "aliphatic1", 80, "white",
  "SL", "100", "Wetting", "aromatic", 30, "black",
  "SL", "100", "Wetting", "aliphatic1", 80, "white"
) %>% 
  dplyr::mutate(sat_level = if_else(treatment=="FM", "FM", as.character(sat_level)),
                sat_level = factor(sat_level, levels = 
                                     c("FM", "5", "35", "50", "75", "100"))) 


ggplot(rel_abund2,
       aes(x = treatment, y = rel_abund, fill = group))+
  geom_bar(stat = "identity")+
  geom_text(data = nmr_label, 
            aes(x = treatment, y = rel_abund, color = color), 
            label = "*", size=7,
            show.legend = F)+
  facet_grid(texture ~ sat_level, scales = "free_x", space = "free_x")+
  scale_x_discrete(position = "bottom") +
  labs(x = "",
       y = "relative contribution (%)",
       subtitle = "% saturation")+
#  scale_fill_viridis_d()+
#  scale_fill_manual(values = soilpalettes::soil_palette("paleustalf",5))+
#  scale_fill_manual(values = PNWColors::pnw_palette("Sunset",5))+
  scale_fill_manual(values = PNWColors::pnw_palette("Sailboat",5))+
  scale_color_manual(values = c("black", "white"))+
  theme_kp()+
  theme(
    plot.subtitle = element_text(size=14, face = "bold", hjust=0.5),
    panel.border = element_rect(color="black",size=0.5, fill = NA),
    strip.background = element_blank(), #facet formatting
    panel.spacing.x = unit(0.5, "lines"), #facet spacing for x axis
    panel.spacing.y = unit(1, "lines"), #facet spacing for x axis
    strip.text.x = element_text(size=12, face = "bold"), #facet labels
    strip.text.y = element_text(size=12, face = "bold", angle = 270)
    )
```

---

### Session Info

Run: `r Sys.Date()`

```{r}
sessionInfo()
```



