---
title: "Hysteresis and Soil C"
author: "Kaizad Patel"
output: github_document
---
Run: `r Sys.Date()`

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE,
                      collapse = TRUE,
                      comment = "#>" #, fig.path = "images/"
                      )

source("3c-picarro_output_NODRAKE.R")
make(plan)

cum_flux = readd(cum_flux)
gf = readd(gf)
meanflux = readd(meanflux)
mean_percsat = readd(mean_percsat)
```


# 1. PICARRO DATA

## negative fluxes removed
## initial plots, QA, QC

1. number of readings


```{r num}
ggplot(cum_flux, aes(x = Core, y = n))+
    geom_point()+
    ggtitle("no. of readings")
```
### NOTE: most readings removed for cores 95, 100, and some others

2. plotting by Core_assignment

```{r, fig.width=8, fig.height=8}
ggplot(gf, aes(DATETIME, flux_co2_umol_g_s*1000, color = Core_assignment)) + 
    geom_point() + geom_line() +
    ylab("flux_co2_nmol_g_s")+
    facet_wrap(~Core_assignment, scale = "free_x")+
    geom_hline(yintercept = 0)+
    theme(legend.position="none")
```

3. each core

```{r cores1, fig.width=15, fig.height=15, include = FALSE}
ggplot(gf, aes(DATETIME, flux_co2_umol_g_s*1000, color = Sand)) + 
    geom_point() + geom_line() +
    ylab("flux_co2_nmol_g_s")+
    facet_wrap(~Core, scale = "free_x")+
    geom_hline(yintercept = 0)+
    theme(legend.position="none")
```

checking individual flux values for each core


```{r cores2}
ggplot(gf, aes(x = as.factor(Core), y = flux_co2_umol_g_s*1000, color = Status, shape = Sand)) + 
  geom_point() +
  scale_shape_manual(values = c(1,19))+
  ylab("flux_co2_nmol_g_s")+
  facet_wrap(~moisture_lvl, scale = "free_x")+
  theme(axis.text.x = element_text(angle=90,hjust=1))
```

5. finding outliers

outlier > 4 SD from the mean for each group

```{r, fig.height=10, fig.width=12}
gf %>% 
  dplyr::select(Core, Core_assignment,Moisture_perc, Sand, Status, soil_type, moisture_lvl,trt,
                DATETIME, flux_co2_umol_g_s, flux_co2_umol_gC_s) %>% 
  group_by(Core_assignment) %>% 
  dplyr::mutate(mean = mean(flux_co2_umol_g_s),
                median = median(flux_co2_umol_g_s),
                sd = sd(flux_co2_umol_g_s),
                se = sd/sqrt(n())) %>% 
  ungroup %>% 
  dplyr::mutate(outlier = if_else((flux_co2_umol_g_s - mean) > 4*sd,"y",as.character(NA)))->gf_test



ggplot(gf_test, aes(x = as.character(Core),y = flux_co2_umol_g_s*1000, color = outlier, shape = soil_type))+
  geom_point()+
  ylab("flux_co2_nmol_g_s")+
  facet_wrap(~Core_assignment, scale = "free")
```



6. checking temporal CV (coefficient of variation) for each core


```{r}
ggplot(cum_flux, aes(x = as.factor(Core), y = cv, color = Status, shape = soil_type))+
  geom_point()+
  scale_shape_manual(values = c(1,19))+
  facet_wrap(~moisture_lvl, scale = "free_x")+
  theme(axis.text.x = element_text(angle=90,hjust=1))

```

---


7. comparing mean and median


```{r}
ggplot(cum_flux, aes(x = as.factor(Core), y = (mean-median)*1000, color = Status, shape = soil_type))+
  geom_point()+
  scale_shape_manual(values = c(1,19))+
  ylab("mean-median, co2_nmol_g_s")+
  facet_wrap(~moisture_lvl, scale = "free_x")+
  theme(axis.text.x = element_text(angle=90,hjust=1))

```


## normalized to soil

```{r soil}
ggplot(cum_flux, aes(x = moisture_lvl, y = cum*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = meanflux, aes(x = as.numeric(moisture_lvl), y = cum*1000), linetype = "longdash", size = 0.5)+
    geom_vline(xintercept = 6.5)+
    ylab("cum flux_co2_nmol_g_s")+
    facet_grid(soil_type~.)+
    ggtitle("cumulative CO2 flux")
  
ggplot(cum_flux, aes(x = moisture_lvl, y = max*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = meanflux, aes(x = as.numeric(moisture_lvl), y = max*1000), linetype = "longdash", size = 0.5)+
    geom_vline(xintercept = 6.5)+
    ylab("maximum flux_co2_nmol_g_s")+
    facet_grid(soil_type~.)+
    ggtitle("maximum CO2 flux")

ggplot(cum_flux, aes(x = moisture_lvl, y = mean*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = meanflux, aes(x = as.numeric(moisture_lvl), y = mean*1000), linetype = "longdash", size = 0.5)+
    geom_vline(xintercept = 6.5)+
    ylab("mean flux_co2_nmol_g_s")+
    facet_grid(soil_type~.)+
    ggtitle("mean CO2 flux")

ggplot(cum_flux, aes(x = moisture_lvl, y = median*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = meanflux, aes(x = as.numeric(moisture_lvl), y = median*1000), linetype = "longdash", size = 0.5)+
    geom_vline(xintercept = 6.5)+
    ylab("mean flux_co2_nmol_g_s")+
    facet_grid(soil_type~.)+
    ggtitle("mean CO2 flux")
```

## normalized to TC content
```{r carbon}
ggplot(cum_flux, aes(x = moisture_lvl, y = cumC*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = meanflux, aes(x = as.numeric(moisture_lvl), y = cumC*1000), linetype = "longdash", size = 0.5)+
    geom_vline(xintercept = 6.5)+
    ylab("cum flux_co2_nmol_gC_s")+
    facet_grid(soil_type~.)+
    ggtitle("cumulative CO2 flux per g C")
  
ggplot(cum_flux, aes(x = moisture_lvl, y = maxC*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = meanflux, aes(x = as.numeric(moisture_lvl), y = maxC*1000), linetype = "longdash", size=0.5)+
    geom_vline(xintercept = 6.5)+
    ylab("maximum flux_co2_nmol_gC_s")+
    facet_grid(soil_type~.)+
    ggtitle("maximum CO2 flux - C")
```

## using percent saturation
```{r perc_sat}

ggplot(cum_flux, aes(x = perc_sat, y = meanC*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = mean_percsat, aes(x = as.numeric(perc_sat), y = meanC*1000), linetype = "longdash", size = 0.5)+
   # geom_vline(xintercept = 6.5)+
    ylab("mean flux_co2_nmol_gC_s")+
    facet_grid(soil_type~.)+
    scale_x_reverse(name = "percent saturation")+
    ggtitle("mean CO2 flux per g C")
  
ggplot(cum_flux, aes(x = perc_sat, y = maxC*1000, color = trt))+
    geom_point(position = position_dodge(width = 0.5))+
    geom_smooth(data = mean_percsat, aes(x = as.numeric(perc_sat), y = maxC*1000), linetype = "longdash", size = 0.5)+
    # geom_vline(xintercept = 6.5)+
    ylab("max flux_co2_nmol_gC_s")+
    facet_grid(soil_type~.)+
    scale_x_reverse(name = "percent saturation")+
    ggtitle("max CO2 flux per g C")
```




## questions
1. Do we delete flux values below zero?

2. Fluxes keep oscillating (Fig. 1). Do we know what's causing that? It's not day-night cycles, since each core was analyzed for only ~24 hours.

3. Soil_50_W (cores 111-115 in Fig. 2) had an initial spike. The samples were initially air-dry and then wetted, so it could be the Birch effect. But this was very short-lived, and this wasn't seen in any of the other "_W" soils. So maybe trailing CO2 from previous samples?

    Do we delete the initial pulse?

4. Soil_sand_dry_W, FIg. 1 (core 95, Fig. 2) had an initial spike and then almost no respiration. Pretty sure that's "wrong" data, since the samples were air-dry and no moisture added. Delete that high value?

5. There were unequal number of readings across cores (Fig. 4), as some ran a little longer than others, like 24 vs. 25 hours. That would influence the cumulative flux calculation. Do we (a) somehow filter data to keep it consistent; (b) leave as is; or (c) represent data a different way (e.g. max/mean flux)?